---
title: "Simulation"
format: html
execute:
  echo: true
  warning: false
  
author: Harper Noteboom
published: 2025/04/01
---

### Simulation

```{r results='hide', warning=FALSE, message=FALSE}
library(tidyverse)

measles <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2020/2020-02-25/measles.csv')

tuesdata <- tidytuesdayR::tt_load('2020-02-25')
tuesdata <- tidytuesdayR::tt_load(2020, week = 9)

measles <- tuesdata$measles

```

For this project I have conducted a permutation test on vaccine rates among different types of schools. The dataset I am using is compiled of 46,412 schools across 32 US States for the 2017-2018 or 2018-2019 (depending on state) school year. This data looks specifically at the vaccine for Measles, Mumps, and Rubella (mmr). The vaccine is federally mandated in public schools but some children are exempt due to medical or religious reasons and there have been other ways that schools can avoid vaccine mandates. I wanted to see if the type of school influenced vaccine rates, the types of schools in this dataset are:

-   Public
-   Private
-   Charter
-   Kindergarten
-   BOCES

Public schools are those funded by the government. Private schools are privately owned and not government funded. Charter schools are also publicly funded but operate outside of the established state school system. Kindergarten are schools for only elementary school students. BOCES is a special program in New York that serves as an extension of publicly funded schools. Despite being publicly funded, charter schools have more flexibility than regular public schools when it comes to curriculum and operations. I wanted to see if this flexibility made a difference in vaccination rates.

My research question is:

In the population, do public schools have higher vaccination rates than charter schools on average?

$H_0$ = $\mu_{\text{public}} = \mu_{\text{charter}}$\
$H_A$ = $\mu_{\text{public}} > \mu_{\text{charter}}$

The graph below shows the average vaccination rate for Public and Charter Schools.

```{r}
c_p_measles <- measles |> 
  select(type, mmr) |> 
  filter(type == "Charter" | type == "Public") |>
  group_by(type) |>
  filter(mmr > 1)|>
  summarize(ave_rate = mean(mmr))

ggplot(c_p_measles, aes(x = type, y = ave_rate, fill = type)) +
  geom_col()+ 
  labs(
    x = "School Type", 
    y = "Average Vaccination Rate", 
    title = "MMR Vaccination Rate at Different Schools"
  )
```

```{r}
c_p_measles |> 
  summarize(ave_diff = diff(ave_rate))
```

I found that the difference between the average vaccine rates for Public and Charter schools to be 8.211511. To test if this is outside the null distribution, I conducted a permutation test in which I scrambled the mmr rates. The histogram of the null distribution can be seen below and the red line marks the observed difference between vaccination rates.

```{r}
perm_test <- function(rep, data) {
  data |>
    select(type, mmr) |> 
    filter(type == "Public" | type == "Charter") |> 
    filter(mmr > 1) |>
    mutate(mmr_perm = sample(mmr, replace = FALSE)) |>
    group_by(type) |>
    summarize(ave_rate = mean(mmr), 
              ave_rate_perm = mean(mmr_perm)) |> 
    summarize(ave_diff = diff(ave_rate), 
              ave_diff_perm = diff(ave_rate_perm), 
              rep = rep)}

set.seed(44)
perm_stats <- 
  map(c(1:500), perm_test, data = measles) |> 
  list_rbind() 

perm_stats |> 
  ggplot(aes(x = ave_diff_perm)) + 
  geom_histogram(bins = 50) + 
  geom_vline(aes(xintercept = ave_diff), color = "red") + 
  labs(
    title = "Null distribution", 
    x = "Average difference"
  )

perm_stats |> 
  summarize(p_val_ave = mean(ave_diff_perm > ave_diff))

```

I found that the p-value is 0 which means the result is statistically significant as p-value \< 0.05. Therefore we can accept the alternative hypothesis and say that on average,public schools have higher vaccination rates than charter schools.

### References

Tidy Tuesday source:\
https://github.com/rfordatascience/tidytuesday/tree/main/data/2020/2020-02-25

Original source: https://github.com/WSJ/measles-data
