---
title: "National Park Species"
format: html
execute:
  echo: false
  warning: false
author: Harper Noteboom
published: 2025/02/19
---

This data comes from the National Parks Service and has information on every species present in national parks across the country. This data set includes data from the 15 most popular parks. The data set contains information on each species' taxonomy and occurrences of the species in the park.

```{r}
#| code-fold: true
#| code-summary: "R Packages"
library(tidyverse)
```

```{r}

#| code-fold: true
#| code-summary: "Filtering and Plot Creation"
#| fig-alt: "Bar graph of the 7 most popular categories of species in Rocky Mountain national park. The x axis has categories of species, in order of highest to lowest number of species : Vasular Plants, Insects, Non-Vascular Plants, Fungi, Birds, Chromista and Mammal. The y axis has number of species and a range from 0 to 1200. There are significantly more species of Vasular Plants than any other category."

most_visited_nps_species_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-10-08/most_visited_nps_species_data.csv', show_col_types = FALSE)

most_visited_nps_species_data |> 
  filter(ParkCode == "ROMO") |> 
  group_by(CategoryName) |> 
  summarize(count = n()) |> 
  filter(count > 50) |> 
  ggplot(aes(x = reorder(CategoryName, -count), y = count, fill = CategoryName)) + 
    geom_col() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1))+
    labs(
      x = "Species Category", 
      y = "Number of species",
      title = "Categories of Species in Rocky Mountain National Park",
      fill = "Category"
    )

```

The figure above shows the number of individual species in different species categories present in Rocky Mountain National Park.

## References:

Original Source: https://irma.nps.gov/NPSpecies/Search/SpeciesList

TidyTuesday Source: https://github.com/rfordatascience/tidytuesday/blob/main/data/2024/2024-10-08/readme.md
