---
title: "Full Text Analysis"
format: html
execute:
  echo: true
  warning: false
  
author: Harper Noteboom
published: 2025/03/05
---

### Dataset

Using a database from BoardGameGeek, a website for board game enthusiasts that features reviews, forums, and board game data, I conducted a full text analysis on games realsed between 1950 and 2016. The dataset has information about each games category, release date, description, average rating, and more.

```{r results='hide', warning=FALSE, message=FALSE}
#| code-fold: true
#| code-summary: "R Packages and Data Loading"

library(tidyverse)

board_games <- readr::read_csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2019/2019-03-12/board_games.csv")

```

### Analysis 1

I first wanted to find out what the most common words in the titles of games were. I filtered out filler words like "the", "and", "to", etc. The top ten most common words board game titles can be seen in the table below.

```{r}
#| code-fold: true
#| code-summary: "Tibble of Most Popular Words"

board_games |>
  mutate(words = str_extract_all(name, "\\b\\w+\\b")) |>  
  unnest(words) |>                                         
  count(words, sort = TRUE) |> 
  filter(!str_detect(words, "^(?i)(the|of|to|game|s|in|and|a|for|at)$"))|>
  head(10)
```

### Analysis 2

Since War was the most common word in game names I decided to look at how many war related games there were. The dataset has a column called 'category' in which the content category of the game is represented. One game can have several categories listed, categories include: fantasy, exploration, adventure, wargame, and many more. I wanted to explore how many games per year are categorized as wargames and if there has been a change in the number of war games released per year. The graphs below explores this question.

```{r}
#| code-fold: true
#| code-summary: "Plot creation - number of war games published per year"
#| fig-alt: "Line graph representing the number of war games published per year from 1955 to 2016, with year on the x axis and count on the y axis. There is a sharp increase from 1970 to 1980, sharp variation from 1980 to 2000 and then a dramatic increase between 2000 and 2010."


war_games <- board_games |> 
  select(name, year_published, category, average_rating) |> 
  group_by(year_published)|> 
  summarize(
    games_per_year = n(), 
    war_games = sum(str_detect(category, "(?i)wargame")), 
    name = name)|> 
  filter(war_games > 0) |> 
  mutate(proportion = war_games/games_per_year)

ggplot(war_games, aes(x = year_published, y = war_games)) + 
geom_point() + 
geom_line() + 
labs(
  title = "Number of war games published per year from 1955-2016", 
  x = "year", 
  y = "number of games published"
)
```

```{r}
#| code-fold: true
#| code-summary: "Plot creation - proportion of war games published per year"
#| fig-alt: "Line graph representing the proportion of war games published per year from 1955 to 2016, with year on the x axis and count on the y axis. There is a decrease from 1960 to 1970 and then a sharp increase until 1980 where it falls until 2000 and increases slightly in 2010."

ggplot(war_games, aes(x = year_published, y = proportion)) + 
geom_point() + 
geom_line() + 
labs(
  title = "Proportion of war games published per year from 1955-2016", 
  x = "year", 
  y = "proportion of total games"
)
```

The first graph shows a sharp increase in the number of war games released between 1970 and 1980 and another sharp increase after 2000. From this graph it would be tempting to say that overall there has been an overall increase in the number of war related game sinse the mid 1900's. However, it is important to take into account the total number of released games. The second graph shows the proportion of games that are considered war games. In this graph we can see that following a low period in the mid-60s, there was a dramatic increase in proportion from the 1970-1980. This aligns with what the first graph showed. The second graph shows that between 1990 and 2000 the proportion of war games decreased and in 2016 the proportion of war games was relatively the same as in 1965. These graphs demonstrate the importance of looking at proportion instead of just quantity.

### Analysis 3

```{r}
#| code-fold: true
#| code-summary: "Tibble of war and average rating"

board_games |> 
  select(name, year_published, category, average_rating) |>
  mutate(war = str_extract(category, "((?<=World War )I+)|(?i)Fantasy.*wargame")) |> 
  mutate(war = str_sub(war, 1, 7))|>
  filter(!is.na(war)) |>
  group_by(war) |> 
  summarize(
    count = n(), 
    average_rating = sum(average_rating)/n())
```

The table above shows how many games were published that were categorized as games about World War I, World War II or Fantasy War. There were almost 6 times as many games made about World War II than World War I and Fantasy War games. Across the three, World War I games have a slightly better average rating but overall they have similar ratings.

### References

Original Source: https://boardgamegeek.com/ Tidy Tuesday Source: https://github.com/rfordatascience/tidytuesday/blob/main/data/2019/2019-03-12/readme.md
